---
title: "STAT 331 Portfolio"
author: "Skylar Richards"
format: 
  html: 
    self-contained: true
layout: margin-left
editor: visual
execute: 
  eval: false
  echo: true
---

[**My Grade:**]{.underline} I believe my grade equivalent to course work evidenced below to be an B

[**Learning Objective Evidence:**]{.underline} In the code chunks below, provide code from Lab or Challenge assignments where you believe you have demonstrated proficiency with the specified learning target. Be sure to specify **where** the code came from (e.g., Lab 4 Question 2).

## Working with Data

**WD-1: I can import data from a *variety* of formats (e.g., csv, xlsx, txt, etc.).**

-   `csv`

```{r}
#| label: wd-1-csv

# from lab 3 
data <- read_csv(here("data","teacher_evals.csv"))


```

-   `xlsx`

```{r}
#| label: wd-1-xlsx
# from practice activity 4
military <- read_xlsx(here::here("data", 
                                 "gov_spending_per_capita.xlsx"), 
                      sheet = "Share of Govt. spending", 
                      skip  = 6, 
                      n_max = 192)


```

-   `txt`

```{r}
#| label: wd-1-txt
# from check in 2.3 
ages_tab <- read_table(file = here::here("Week 2", "Check-ins", "Ages_Data", "ages_tab.txt"))


```

**WD-2: I can select necessary columns from a dataset.**

```{r}
#| label: wd-2
# from lab 2, Q6

teacher_evals_clean |>
  filter(if_any(everything(), ~is.na(.))) |>
  select(course_id,teacher_id, everything())


```

**WD-3: I can filter rows from a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   numeric

```{r}
#| label: wd-3-numeric

# from lab 5

suspect_killer <- drivers_license|>
  filter(gender == "male",
         str_detect(plate_number, "H42W"))|>
  rename(license_id = id) |>
 inner_join(person, by = "license_id") |>
  inner_join(get_fit_now_member, by = c("id" = "person_id", "name" = "name"))|>
  rename(membership_id = id.y)|>
  inner_join(get_fit_now_check_in, by = "membership_id")|>
  filter(check_in_date == "20180109", # numeric filtering here
         str_detect(membership_id, "^48Z"))


```

-   character -- specifically a string (example must use functions from **stringr**)

```{r}
#| label: wd-3-string

# from lab 5

suspect_killer <- drivers_license|>
  filter(gender == "male",
         str_detect(plate_number, "H42W"))|>
  rename(license_id = id) |>
 inner_join(person, by = "license_id") |>
  inner_join(get_fit_now_member, by = c("id" = "person_id", "name" = "name"))|>
  rename(membership_id = id.y)|>
  inner_join(get_fit_now_check_in, by = "membership_id")|>
  filter(check_in_date == "20180109",
         str_detect(membership_id, "^48Z")) #string filtering here
```

-   factor

```{r}
#| label: wd-3-factor



```

-   date (example must use functions from **lubridate**)

```{r}
#| label: wd-3-date

# from lab 5

filtered_crime_scene_report <- crime_scene_report|>
  mutate(date = ymd(date))|> # convert dates into date format 
  filter(type == "murder",
         date == "2018-01-15",
         city == "SQL City") 

```

**WD-4: I can modify existing variables and create new variables in a dataframe for a *variety* of data types (e.g., numeric, integer, character, factor, date).**

-   numeric (using `as.numeric()` is not sufficient)

```{r}
#| label: wd-4-numeric

fish_challenge <- fish|>
  mutate(length = length/10)

# from Challenge 7 
```

-   character -- specifically a string (example must use functions from **stringr**)

```{r}
#| label: wd-4-string

ca_childcare <- ca_childcare |> 
  mutate(county_name = str_remove(county_name, " County")) 
  
# from lab 4

```

-   factor (example must use functions from **forcats**)

```{r}
#| label: wd-4-factor

# from lab 4
ca_childcare <- ca_childcare |> 
  mutate(county_name = str_remove(county_name, " County")) |> # removes county and blank space before
  mutate(region = fct_collapse(county_name,  "Superior California" = c("Butte", "Colusa", "El Dorado",
                                                         "Glenn", "Lassen", "Modoc",
                                                         "Nevada", "Placer", "Plumas",
                                                         "Sacramento", "Shasta", "Sierra",
                                                         "Siskiyou", "Sutter", "Tehama",
                                                         "Yolo", "Yuba", ...

                               # fct_collpase involves 20+ lines, shortened for readability 
```

-   date (example must use functions from **lubridate**)

```{r}
#| label: wd-4-date

filtered_crime_scene_report <- crime_scene_report|>
  mutate(date = ymd(date))

# from lab 4

```

**WD-5: I can use mutating joins to combine multiple dataframes.**

-   `left_join()`

```{r}
#| label: wd-5-left

```

-   `right_join()`

```{r}
#| label: wd-5-right

```

-   `inner_join()`

```{r}
#| label: wd-5-inner
# from lab 5

interview_witness <- filtered_person|> #get interviews of witnesses
  inner_join(interview,
             by = c("id" = "person_id"))


```

-   `full_join()`

```{r}
#| label: wd-5-full

```

**WD-6: I can use filtering joins to filter rows from a dataframe.**

-   `semi_join()`

```{r}
#| label: wd-6-semi

```

-   `anti_join()`

```{r}
#| label: wd-6-anti

```

**WD-7: I can pivot dataframes from long to wide and visa versa**

-   `pivot_longer()`

```{r}
#| label: wd-7-long

results_long <- pivot_longer(
  results_table,
  cols = everything(),
  names_to = "ncorrect",
  values_to = "count")

# from lab 9 
```

-   `pivot_wider()`

```{r}
#| label: wd-7-wide

survey_data_type <- surveys|>
  map_chr(class)|>
  enframe(name = "Variable", value = "DataType")|>
  pivot_wider(names_from = Variable, values_from = DataType)

# from lab 8
```

## Reproducibility

**R-1: I can create professional looking, reproducible analyses using RStudio projects, Quarto documents, and the here package.**

I've done this in the following provided assignments:

Lab

**R-2: I can write well documented and tidy code.**

-   Example of **ggplot2** plotting

```{r}
#| label: r-2-1

ggplot(all_simulations, aes(x = simulated_means)) +
  geom_histogram(binwidth = 0.2,
                 fill = "steelblue",
                 color = "black",
                 alpha = 0.7) +
  facet_wrap(~n, scales = "free_y", 
             labeller = as_labeller(c(`10` = "n = 10",
                                      `100` = "n = 100",
                                      
                                      `1000` = "n = 1000",
                                      `10000` = "n = 10000"))) +
  labs(
    title = "Distribution of Simulated Means by Sample Size",
    x = "Simulated Means",
    y = "Frequency"
)

# from lab 9 

```

-   Example of **dplyr** pipeline

```{r}
#| label: r-2-2

```

-   Example of function formatting

```{r}
#| label: r-2-3
rescale_column <- function(data_frame, ...) {
  data_frame |>
    mutate(across(c(...), ~ rescale_01(.)))
}

# From Lab 7 
```

**R-3: I can write robust programs that are resistant to changes in inputs.**

-   Example -- any context

```{r}
#| label: r-3-example

```

-   Example of function stops

```{r}
#| label: r-3-function-stops

```

## Data Visualization & Summarization

**DVS-1: I can create visualizations for a *variety* of variable types (e.g., numeric, character, factor, date)**

-   at least two numeric variables

```{r}
#| label: dvs-1-num
# from lab 4

ggplot(ca_childcare, aes(x = mhi_2018, y = mc_infant)) +
  geom_point(color = "blue", size = 3, alpha = 0.4) + 
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(
    title = "Full Time median weekly price for center-based childcare for an infant in California",
    x= "Median household income (2018 dollars)",
    y="")+
theme_minimal()

```

-   at least one numeric variable and one categorical variable

```{r}
#| label: dvs-2-num-cat
# from lab 4

ggplot(lowest_infant_care_2018, aes(x=median_care, y=reorder(region, -median_care)))+
  geom_col(fill = "steelblue")+
  facet_wrap(~study_year)+
  ggtitle("2018 Infant care cost by region") +
  ylab("") +
  xlab("Median Care cost in 2018 dollars") +
  theme_minimal()

```

-   at least two categorical variables

```{r}
#| label: dvs-2-cat
# from lab 4

ggplot(lowest_infant_care, aes(x=median_care, y=reorder(region, -median_care)))+
  geom_col(fill = "steelblue")+
  facet_wrap(~study_year)+
  ggtitle("2008 vs 2018 Infant care cost by region") +
  ylab("") +
  xlab("Median Care cost in 2018 dollars") +
  theme_minimal()
```

-   dates (timeseries plot)

```{r}
#| label: dvs-2-date

ggplot(median_prices_regional, aes(x=study_year,y=price_care, color = region))+
   geom_point(position = position_dodge(width = 0)
              , alpha = 0.8,
              size = 0.5) +  
  geom_smooth(method = "loess",
              se = TRUE,
              alpha = 0.5)+
  facet_wrap(~ age_group,
             nrow = 1)+
  scale_y_continuous()+
  labs(
  title = "Weekly Median Price for Center-Based Childcare ($)",
  x = "Study Year",
  y = "",
  color = "California Region")

# from lab 4


```

**DVS-2: I use plot modifications to make my visualization clear to the reader.**

-   I can ensure people don't tilt their head

```{r}
#| label: dvs-2-1

```

-   I can modify the text in my plot to be more readable

```{r}
#| label: dvs-2-2

```

-   I can reorder my legend to align with the colors in my plot

```{r}
#| label: dvs-2-3



```

**DVS-3: I show creativity in my visualizations**

-   I can use non-standard colors

```{r}
#| label: dvs-3-1
# from lab 4

ggplot(median_prices_regional, aes(x=study_year,y=price_care, color = region))+
   geom_point(position = position_dodge(width = 0)
              , alpha = 0.8,
              size = 0.5) +  
  geom_smooth(method = "loess",
              se = TRUE,
              alpha = 0.5)+
  facet_wrap(~ age_group,
             nrow = 1)+
  scale_y_continuous()+
  labs(
  title = "Weekly Median Price for Center-Based Childcare ($)",
  x = "Study Year",
  y = "",
  color = "California Region")+
  scale_x_continuous(breaks = seq(2008,2018,2))+
  scale_y_continuous(limits = c(100,500))+
  scale_color_manual(values = c("San Francisco Bay Area" = "green",
                                "Orange County" = "grey",
                                "Los Angeles County" = "pink",
                                "Northern San Joaquin Valley" = "orange",
                                "Central Coast" = "yellow",
                                "Inland Empire" = "blue",
                                "Superior California" = "maroon",
                                "Southern San Joaquin Valley" = "red",
                                "San Diego - Imperial" = "brown",
                                "North Coast" = "darkgrey"))+ ...

```

-   I can use annotations

```{r}
#| label: dvs-3-2



```

-   I can be creative...

```{r}
#| label: dvs-3-3
# from lab 4

ggplot(median_prices_regional, aes(x=study_year,y=price_care, color = region))+
   geom_point(position = position_dodge(width = 0)
              , alpha = 0.8,
              size = 0.5) +  
  geom_smooth(method = "loess",
              se = TRUE,
              alpha = 0.5)+
  facet_wrap(~ age_group,
             nrow = 1)+
  scale_y_continuous()+
  labs(
  title = "Weekly Median Price for Center-Based Childcare ($)",
  x = "Study Year",
  y = "",
  color = "California Region")+
  scale_x_continuous(breaks = seq(2008,2018,2))+
  scale_y_continuous(limits = c(100,500))+
  scale_color_manual(values = c("San Francisco Bay Area" = "green",
                                "Orange County" = "grey",
                                "Los Angeles County" = "pink",
                                "Northern San Joaquin Valley" = "orange",
                                "Central Coast" = "yellow",
                                "Inland Empire" = "blue",
                                "Superior California" = "maroon",
                                "Southern San Joaquin Valley" = "red",
                                "San Diego - Imperial" = "brown",
                                "North Coast" = "darkgrey"))+
  theme(
    axis.text = element_text(size = 6),
    panel.border = element_rect(color = "black", fill = NA),
    aspect.ratio = 1,
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8),
    legend.key.height = unit(-1,"cm"),
    legend.key.size = unit(1,"cm"),
    legend.key.spacing.y = unit(0.1,"cm"),
    panel.background = element_rect(fill = "white", color = NA),
    panel.grid = element_line(color = "grey", size = 0.15),
    strip.background = element_rect(fill = "grey", color = "black", size = 1),
    title = element_text(size = 7.5)
    )
```

**DVS-4: I can calculate numerical summaries of variables.**

-   Example using `summarize()`

```{r}
#| label: dvs-4-summarize
# from lab 4

lowest_infant_care <- ca_childcare|>
  filter(study_year %in% c(2018,2008))|>
  group_by(study_year, region)|>
  summarize(median_care = median(mc_infant), .groups = "drop")|>
  arrange(desc(median_care))
```

-   Example using `across()`

```{r}
#| label: dvs-4-across

missing_values <- fish|>
 summarise(
    total_missing_observations = sum((across(everything(), ~ sum(is.na(.))))),
    variables_missing = names(which(colSums(is.na(fish))>0))
  )
  
# from lab 7 

```

**DVS-5: I can find summaries of variables across multiple groups.**

-   Example 1

```{r}
#| label: dvs-5-1

missing_values_fish <- fish|>
  map_int(~sum(is.na(.)))|>
  enframe(name = "Variable Name", value = "Number of Missing Values")

# from lab 8

```

-   Example 2

```{r}
#| label: dvs-5-2
summary_table <- evals|>
  group_by(teacher_id)|>
  summarize(
    sex = first(sex),
    seniority = mean(seniority, na.rm = TRUE),
    academic_degree = first(academic_degree),
    .groups = "drop"
  )|>
  mutate(seniority_level = if_else(seniority <= 4, "Junior", "Senior"))|>
  count(sex,seniority_level, academic_degree)|>
  pivot_wider(names_from = c(sex, seniority_level, academic_degree), values_from = n)|>
  summarize(
    Male = sum(c(male_Junior_dr,
                 male_Junior_ma,
                 male_Junior_no_dgr,
                 male_Junior_prof,
                 male_Senior_dr,
                 male_Senior_ma,
                 male_Senior_no_dgr,
                 male_Senior_prof), na.rm = TRUE),
  Female = sum(c(female_Junior_dr,
                 female_Junior_ma,
                 female_Junior_no_dgr,
                 female_Junior_prof,
                 female_Senior_dr,
                 female_Senior_ma), na.rm = TRUE),
  "Junior (Less than 4 years)"  = sum(c(male_Junior_dr,
                 male_Junior_ma,
                 male_Junior_no_dgr,
                 male_Junior_prof,
                 female_Junior_dr,
                 female_Junior_ma,
                 female_Junior_no_dgr,
                 female_Junior_prof), na.rm = TRUE),
   "Senior (More than 4 years)" = sum(c(male_Senior_dr,
                 male_Senior_ma,
                 male_Senior_no_dgr,
                 male_Senior_prof,
                 female_Senior_dr,
                 female_Senior_ma), na.rm = TRUE),
  "No Degree" = sum(c(male_Senior_no_dgr,
                    male_Junior_no_dgr,
                    female_Junior_no_dgr), na.rm = TRUE),
  Masters = sum(c(male_Senior_ma,
                    male_Junior_ma,
                    female_Junior_ma,
                    female_Senior_ma), na.rm = TRUE),
  Doctorate = sum(c(male_Senior_dr,
                    male_Junior_dr,
                    female_Junior_dr,
                    female_Senior_dr), na.rm = TRUE),
  Professor = sum(c(male_Senior_prof,
                    male_Junior_prof,
                    female_Junior_prof), na.rm = TRUE)
  )
# from lab 8 
```

**DVS-6: I can create tables which make my summaries clear to the reader.**

-   Example 1

```{r}
#| label: dvs-6-1
# from lab 4

combined_lowest <-bind_rows(
  lowest_2018,
  lowest_2008)|>
  rename("Lowest Median Regional Price for Infant care in California" = "median_care",
         "Year" = "study_year")
print(combined_lowest)

```

-   Example 2

```{r}
#| label: dvs-6-2
pretty_summary_table_gt <- summary_table |>
  gt() |>
  tab_header(
    title = "Instructor Evaluations Demographics"
  ) |>
  tab_style(
    style = cell_fill(color = "lightgray"),
    locations = cells_body(rows = seq(1, nrow(summary_table), 2))
  ) |>
  tab_options(
    table.font.size = px(14),
    table.border.top.color = "black",
    table.border.top.width = px(2),
    table.border.bottom.color = "black",
    table.border.bottom.width = px(2),
    table.border.left.color = "black",
    table.border.left.width = px(2),
    table.border.right.color = "black",
    table.border.right.width = px(2),  
    row.striping.include_table_body = TRUE,  
    row.striping.background_color = "lightgray", 
    data_row.padding = px(5),  
    column_labels.font.size = px(14), 
    column_labels.font.weight = "bold",
    column_labels.background.color = "lightblue", 
    table.width = pct(100)  
  )

# from lab 9 
```

**DVS-7: I show creativity in my tables.**

-   Example 1

```{r}
#| label: dvs-7-1

pretty_summary_table_gt <- summary_table |>
  gt() |>
  tab_header(
    title = "Instructor Evaluations Demographics"
  ) |>
  tab_style(
    style = cell_fill(color = "lightgray"),
    locations = cells_body(rows = seq(1, nrow(summary_table), 2))
  ) |>
  tab_options(
    table.font.size = px(14),
    table.border.top.color = "black",
    table.border.top.width = px(2),
    table.border.bottom.color = "black",
    table.border.bottom.width = px(2),
    table.border.left.color = "black",
    table.border.left.width = px(2),
    table.border.right.color = "black",
    table.border.right.width = px(2),  
    row.striping.include_table_body = TRUE,  
    row.striping.background_color = "lightgray", 
    data_row.padding = px(5),  
    column_labels.font.size = px(14), 
    column_labels.font.weight = "bold",
    column_labels.background.color = "lightblue", 
    table.width = pct(100)  
  )

# from lab 9 
```

-   Example 2

```{r}
#| label: dvs-7-2
pretty_summary_table_kable <- summary_table|>
   kable(caption = "Instructor Evaluations Demographics") |>
   kable_styling(bootstrap_options = c("striped", "hover", "condensed", "bordered"), 
                 full_width = TRUE, 
                 font_size = 14)

# from lab 9 
```

## Program Efficiency

**PE-1: I can write concise code which does not repeat itself.**

-   using a single function call with multiple inputs (rather than multiple function calls)

```{r}
#| label: pe-1-one-call

rescale_column <- function(data_frame, ...) {
  data_frame |>
    mutate(across(c(...), ~ rescale_01(.)))
}

# from lab 7
```

-   `across()`

```{r}
#| label: pe-1-across
missing_values <- fish|>
 summarise(
    total_missing_observations = sum((across(everything(), ~ sum(is.na(.))))),
    variables_missing = names(which(colSums(is.na(fish))>0))
  )

# lab 7
```

-   `map()` functions

```{r}
#| label: pe-1-map-1

missing_values_fish <- fish|>
  map_int(~sum(is.na(.)))|>
  enframe(name = "Variable Name", value = "Number of Missing Values")

# from lab 8 
```

**PE-2: I can write functions to reduce repetition in my code.**

-   Function that operates on vectors

```{r}
#| label: pe-2-1
rescale_01 <- function(value){
  if ((length(value) > 1) & is.numeric(value)) {
   (value = (value - min(value, na.rm = TRUE)) / 
           (max(value, na.rm = TRUE) - min(value, na.rm = TRUE)))
  }}

# from lab 7 
```

-   Function that operates on data frames

```{r}
#| label: pe-2-2
rescale_column <- function(data_frame, ...) {
  data_frame |>
    mutate(across(c(...), ~ rescale_01(.)))
}

# from lab 7 
```

**PE-3:I can use iteration to reduce repetition in my code.**

-   `across()`

```{r}
#| label: pe-3-across

rescale_column <- function(data_frame, ...) {
  data_frame |>
    mutate(across(c(...), ~ rescale_01(.)))
}

# from lab 7 
```

-   `map()` function with **one** input (e.g., `map()`, `map_chr()`, `map_dbl()`, etc.)

```{r}
#| label: pe-3-map-1

```

-   `map()` function with **more than one** input (e.g., `map_2()` or `pmap()`)

```{r}
#| label: pe-3-map-2

```

**PE-4: I can use modern tools when carrying out my analysis.**

-   I can use functions which are not superseded or deprecated

```{r}
#| label: pe-4-1

```

-   I can connect a data wrangling pipeline into a `ggplot()`

```{r}
#| label: pe-4-2

```

## Data Simulation & Statisical Models

**DSSM-1: I can simulate data from a *variety* of probability models.**

-   Example 1

```{r}
#| label: dsm-1-1

randomBabies <- function(nBabies = 4){
  step1 <- sample(1:nBabies, nBabies, replace = FALSE)
  step2 <- sum(step1 == 1:nBabies)
  return(step2)
  
# from lab 9 
```

-   Example 2

```{r}
#| label: dsm-1-2
simulate_means <- function(n, df){
  map_dbl(.x = 1:n, 
          .f = ~rchisq(n = 100, df =df) |> mean()
          )
}

# from lab 9 
```

**DSSM-2: I can conduct common statistical analyses in R.**

-   Example 1

```{r}
#| label: dsm-2-1
# 
species_mod <- aov()

summary(species_mod)
```

-   Example 2

```{r}
#| label: dsm-2-2

```

## Revising My Thinking

<!-- How did you revise your thinking throughout the course? How did you revise your thinking on the code examples you have provided in your portfolio? -->

One of the biggest ways of revising my thinking was through streamlining operations. I accomplished this mostly through reducing the number of temporary objects found in my code which helps compress operations and keeps code tidy.\

<!-- For the revisions included in your Portfolio, to help me understand the nature of your revisions, please denote somehow the feedback I provided you (e.g., boldface, italics, colored text) before your revisions. -->

## Extending My Thinking

<!-- How did you extended your thinking throughout the course? How did you extend your thinking on the code examples you have provided in your portfolio? -->

The biggest way I have extended my thinking is how I consider data organization to help it be easier to process and/or understand. organizing the data in a way which reduces the number of required transformations or joining of datasets helps the raw data be easier to analyze.

## Peer Support & Collaboration

<!-- Include an image or a description of feedback you gave that you are proud of (either in a peer review or in Discord). -->

Often times during practice activities in class groups nearby will ask for help and instead of providing them the answer directly I will help lead them to finding it themselves.

<!-- Include a description of how you grew as a collaborator through the weekly pair programming activities.   -->

I think the most beneficial aspect is being exposed to different people's way of thinking about the approach to a problem. This helps both in communicating goals/processes to people who think differently and also help expand my own line of thinking.
